---
title: "Titanic"
author: "Chen Shih-Chieh"
date: "July 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This analysis takes a lot of inspiration from the Titanic data set tutorial on kaggle. I've recently gotten into machine learning and this will be for practicing what I've learned


Loading the scripts

```{r, results = "hide", warning=FALSE, message=FALSE}
library(tidyverse) #data maniplulation, visualization, etc
library(caret) #model creation
library(randomForest) #model creation

train_dat <- read.csv("dat/train.csv")
test_dat <- read.csv("dat/test.csv")
```

Taking a look at the data

```{r}
str(train_dat)
```


The names variable seems to be able to be broken down into other lables, e.g. ms, miss, doctor, mr. 
Taking a look at the Names variable. There are a lot of observations so I'm  just going to look at a subset of them.
```{r, results = "hide"}
train_dat$Name[sample(1:nrow(train_dat), 100)]
```

It seems that a persons title can be extracted fairly easily.
```{r}
train_dat$title <- train_dat$Name %>%
  str_match(., ", (([:alpha:]|\\s)*)\\.") %>% .[,2]
```

Taking a look at how many different titles we have. 
```{r}
str(train_dat$title) #there seems to be 18 different titles
table(train_dat$title) #taking a look at the distribution
```

Lets reassign the titles mlle, ms, and mme and group together titles that are rare
```{r}
train_dat$title[train_dat$title == "Mlle"] <- "Miss"
train_dat$title[train_dat$title == "Ms"] <- "Miss"
train_dat$title[train_dat$title == "Mme"] <- "Mrs"

rare_titles <- c("Capt", "Col", "Don", "Dr", "Lady", "Major", "Rev", "Sir", "Jonkheer", "the Countess")
train_dat$title[train_dat$title %in% rare_titles] <- "Rare Title"
```

Out of curiosity, taking a look at the distribution on survival and title

```{r}
train_dat %>% ggplot(aes(title, fill = factor(Survived))) +
  geom_bar(stat = "count", position = "dodge")
```

Next, lets see how family size affects survival
```{r}
#calculating family size
train_dat$fullfamily <- train_dat$SibSp + train_dat$Parch + 1 
train_dat %>% ggplot(aes(fullfamily, fill = factor(Survived))) +
  geom_bar(stat = "count", position = "dodge")
```

Lets group people into 3 parts. Singeton, small family, and large family. First the break down of data points in each group.

```{r}
table(train_dat$Survived, train_dat$fullfamily)
table(train_dat$Survived, train_dat$fullfamily)[2,]/colSums(table(train_dat$Survived, train_dat$fullfamily))
```


Looks there theres a difference between the survival rate if a person is by themselves, with a family of 1 to 3, and with a family of 4 or more. So the family is divided that way.
```{r}
 train_dat$FamilySize[train_dat$fullfamily == 1] <- "Singleton"
 train_dat$FamilySize[train_dat$fullfamily > 1 & train_dat$fullfamily < 5] <- "Small Family"
 train_dat$FamilySize[train_dat$fullfamily >= 5] <- "Large Family"
```

Is there a difference in survivability and what deck someones room is at? Extracting deck letter from the cabin number.
```{r}
train_dat$Deck <- train_dat$Cabin %>% str_match("[:alpha:]")
```

Looking at the data, there are many missing values for Age.
```{r}
colSums(is.na(train_dat))
```

Filling in the missing age values by computing them. Using mice, aka multiple imputation by chained equation. 
