---
title: "Titanic"
author: "Chen Shih-Chieh"
date: "July 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This analysis takes a lot of inspiration from the Titanic data set tutorial on kaggle. I've recently gotten into machine learning and this will be for practicing what I've learned


Loading the scripts

```{r, results = hide}
library(tidyverse) #data maniplulation, visualization, etc
library(caret) #model creation
library(randomForest) #model creation

train_dat <- read.csv("dat/train.csv")
test_dat <- read.csv("dat/test.csv")
```

Taking a look at the data

```{r}
str(train_dat)
```


The names variable seems to be able to be broken down into other lables, e.g. ms, miss, doctor, mr. 
Taking a look at the Names variable. There are a lot of observations so I'm  just going to look at a subset of them.
```{r, results = hide}
train_dat$Name[sample(1:nrow(train_dat), 100)]
```

It seems that a persons title can be extracted fairly easily.
```{r}
train_dat$title <- train_dat$Name %>%
  str_match(., ", ([:alpha:]*)\\.") %>% .[,2] %>% factor
```

Taking a look at how many different titles we have. 
```{r}
str(train_dat$title) #there seems to be 18 different titles
levels(train_dat$title)
table(train_dat$title) #taking a look at the distribution
```

